---
title: "Stat 602 Term Project"
author: "Ethan Scott"
date: "2023-10-04"
output:
  word_document: default
  html_document:
    df_print: paged
toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#### Loading Libraries Needed

```{r}
library(ggplot2)
library(dplyr)
library(glmnet)
library(caret)
library(Metrics)
```

## Reading Data

```{r}
listings = read.csv("https://raw.githubusercontent.com/ethan2411/Data-601-602/main/listings.csv")
head(listings)
colnames(listings)
```

Converting price to be numeric

```{r}
# Converting price to be numeric, removing $ and , to make it possible
listings = listings %>%
  mutate(price = as.numeric(gsub("[$,]", "", price)))

# Check the first few rows of the updated dataframe
head(listings)
```

## Visualizing Price by Neighborhood

```{r}
ggplot(data = listings, aes(x=price, y = neighbourhood_cleansed)) + geom_violin(col="blue")+ geom_boxplot(col="purple") + theme_minimal()
```

```{r}
# Filter out the outliers so it is easier to see the graphs
listings_filtered = listings %>%
  filter(price < 1500)

# Create the plot using the filtered data
ggplot(data = listings_filtered, aes(x = price, y = neighbourhood_cleansed)) +
  geom_boxplot(col = "purple") +
  theme_minimal() +
  labs(
    title = "Distribution of Price by Neighbourhood",
    x = "Price",
    y = "Neighbourhood"
  )
```

```{r}
# Create the plot using the filtered data
ggplot(data = listings_filtered, aes(x = price, y = neighbourhood_cleansed)) +
  geom_violin(col = "purple") +
  theme_minimal() +
  labs(
    title = "Distribution of Price by Neighbourhood",
    x = "Price",
    y = "Neighbourhood"
  )

```

## Visualizing Price by Room Type

```{r}
# Create the plot using the filtered data
ggplot(data = listings_filtered, aes(x = price, y = room_type)) +
  geom_violin(col = "blue", width= 1.5)+
  geom_boxplot(col = "purple", width=0.2) +
  theme_minimal() +
  labs(
    title = "Distribution of Price by Property Type",
    x = "Price",
    y = "Property Type"
  )
```

## Confidence Intervals

```{r}
#creating the confidence intervals for each of the listings
neighborhood_means = listings %>%
  group_by(neighbourhood_cleansed) %>%
  summarize(
    mean_price = mean(price, na.rm = TRUE),
    lower_ci = t.test(price, conf.level = 0.95)$conf.int[1],
    upper_ci = t.test(price, conf.level = 0.95)$conf.int[2]
  )
head(neighborhood_means)
#plotting the confidence intervals
ggplot(neighborhood_means, aes(x = reorder(neighbourhood_cleansed, mean_price), y = mean_price)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower_ci, ymax = upper_ci), width = 0.2) +
  coord_flip() +
  labs(
    x = "Neighborhood",
    y = "Mean Price",
    title = "Mean Price by Neighborhood with Confidence Intervals"
  ) +
  theme_minimal() +
  theme(axis.text.y = element_text(size = 10))
```









## Regression

### Data Cleaning

```{r}
# Seeing what columns have too many missing values, we will remove these
missing_values = is.na(listings)
missing_counts = colSums(missing_values)
columns5000_missing = missing_counts[missing_counts >5000]
print(columns5000_missing)
```

#### Feature Engineering

```{r}
#create a feature for if a host has an about section? remove host_about
#other features too probably

#maybe how long someone has been a host will affect price?
listings$host_since = as.Date(listings$host_since)
listings$hostfor = as.numeric(difftime(Sys.Date(), listings$host_since, units = "days"))

sum(is.na(listings$hostfor))
# fill missing values in 'hostfor' with the mean
mean_hostfor = mean(listings$hostfor, na.rm = TRUE)
listings$hostfor[is.na(listings$hostfor)] = mean_hostfor
sum(is.na(listings$hostfor))
```

```{r}
# fill missing 'beds' values with the median 
# median is 2 and mean is 2.1 something
median_beds = median(listings$beds, na.rm = TRUE)
listings$beds[is.na(listings$beds)] = median_beds
```

```{r}
#maybe if the host includes an in depth summary about themselves it will affect price?
listings$host_about_word_count = sapply(listings$host_about, function(text) {
  words = unlist(strsplit(text, " "))
  return(length(words))
})

#same for a description of the place?

```

```{r}
# Extract the number of bathrooms available
# Need to find out how to also carry over decimal and "half-bath" listings
listings$bathrooms_count = as.numeric(gsub("[^0-9]", "", listings$bathrooms_text))


print(unique(listings$bathrooms_count))
print(unique(listings$bathrooms_text))
```

#### Removing Unnecessary Columns

```{r}
#removing columns that will definitely have nothing to do with price
listings = listings[, !(colnames(listings) %in% c("listing_url", "scrape_id", "last_scraped", "source", "picture_url", "host_id", "host_url", "host_thumbnail_url", "host_picture_url", "neighbourhood", "neighbourhood_group_cleansed", "license", "calendar_last_scraped", "calendar_updated", "property_type", "bedrooms", "host_since", "host_location", "host_about", "bathrooms_test", "bathrooms", "review_scores_communication", "review_scores_location", "review_scores_value", "reviews_per_month", "review_scores_accuracy", "review_scores_cleanliness", "review_scores_checkin", "host_listings_count", "host_total_listings_count", "last_review", "host_name", "calculated_host_listings_count", "calculated_host_listings_count_entire_homes", "calculated_host_listings_count_private_rooms", "calculated_host_listings_count_shared_rooms"))] 

#bathrooms was removed at end because only bathroom_text will be needed, will probably have to do some feature engineering for lots of these variables
listings = listings[, !(colnames(listings) %in% c("description", "neighborhood_overview", "latitude", "longitude", "name", "amenities", "id"))] 

head(listings)
colnames(listings)
#head(listings$beds)
```

```{r}
sapply(listings, function(x) length(unique(x)))
```

```{r}
#head(listings$calculated_host_listings_count_entire_homes)
```

#### Regression Modelling

```{r}
# Split the data into training and testing sets
set.seed(123)  # For reproducibility
options(scipen = 999) # For Readability
trainIndex = createDataPartition(listings$price, p = .8, 
                                  list = FALSE, 
                                  times = 1)
data_train = listings[ trainIndex,]
data_test  = listings[-trainIndex,]

# Fit a linear regression model  
#WILL need to change the x variables later, once other features are decided
model = lm(price ~beds+hostfor + accommodates, data = data_train)

# Make predictions on the test data
predictions = predict(model, newdata = data_test)
#summary of the model
summary(model)

rmse_value = rmse(predictions, data_test$price)
mse_value = mse(predictions, data_test$price)
mae_value = mae(predictions, data_test$price)

# Print the evaluation metrics
cat("Mean Squared Error (MSE):", mse_value, "\n")
cat("Root Mean Squared Error (RMSE):", rmse_value, "\n")
cat("Mean Absolute Error (MAE):", mae_value, "\n")
```

#### Ridge/Lasso/Elastic Net Regression

```{r}
#this is the code from a previous example so will definitely need to adjust it a bit
#lasso has alpha =1 and elastic net anything between 0 and 1, most commonly 0.5


#cv.ridge = cv.glmnet(x.ridge, y.ridge, alpha = 0, family = "binomial")
#model.ridge = glmnet(x.ridge, y.ridge, alpha =0, family = "binomial", lambda = cv.ridge$lambda.min)
#x.ridge.test = model.matrix(TenYearCHD ~., test_data)[,-1]
#probabilities.ridge = model.ridge %>% predict(newx = x.ridge.test)
```



### Anova testing (james)

H0: There is no 

```{r}
options(scipen = 999)
mean.aov = aov(price ~ accommodates*beds*bathrooms_count*number_of_reviews, data = listings)

summary(mean.aov)
```
```{r}
options(scipen = 999)
mean.aov = aov(review_scores_rating ~ accommodates+bathrooms_count+price, data = listings)

summary(mean.aov)
```
```{r}
plot(mean.aov)
```



## Conclusions

stuff here

## Citations

```{r echo=TRUE}
citation('ggplot2')
citation('dplyr')
citation('glmnet')
citation('caret')
citation('Metrics')
```
